---
title: "Analysis of Submetered Household Electric Power Consumption"
author: "Ashiqur"
date: "11/20/2019"
output: html_document
---
```{r setup, include=TRUE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

```{r library}
library(RMySQL)
library(dplyr)
library(lubridate)
library(VIM)        #Aids visualization and imputing of missing values
library(funModeling)
library(ggplot2)
library(reshape2)
library(caret)      #R modeling workhorse & ggplot2
library(tidyverse)  #Package for tidying datalibrary(magrittr)   #Enables piping
library(GGally)
```

## Data Connection & Retrieving Data
##### Create a database connection 
```{r load_data}
con = dbConnect(MySQL(), user='deepAnalytics', 
                password='Sqltask1234!', 
                dbname='dataanalytics2018', 
                host='data-analytics-2018.cbrosir2cswx.us-east-1.rds.amazonaws.com')
```

### List the tables contained in the database
```{r table}
dbListTables(con)
```
### Lists attributes contained in a table
```{r attributes}
dbListFields(con,'yr_2006')
```

## Request relevant data base tables

#### Use asterisk to specify all attributes for download

```{r pressure}
yr_2006 <- dbGetQuery(con, "SELECT * FROM yr_2006")
yr_2007 <- dbGetQuery(con, "SELECT * FROM yr_2007")
yr_2008 <- dbGetQuery(con, "SELECT * FROM yr_2008")
yr_2009 <- dbGetQuery(con, "SELECT * FROM yr_2009")
yr_2010 <- dbGetQuery(con, "SELECT * FROM yr_2010")
```
## Investigate each new data frame

#### Year 2006
```{r y_2006}
summary(yr_2006)
str(yr_2006)
head(yr_2006)
tail(yr_2006)
```
The year 2006 doesn't have the entire year Data. It has only 15 days of data.

#### Year 2007
```{r y_2007}
summary(yr_2007)
str(yr_2007)
head(yr_2007)
tail(yr_2007)
```
The year 2007 has entire year Data

#### Year 2008
```{r y_2008}
summary(yr_2008)
str(yr_2008)
head(yr_2008)
tail(yr_2008)
```
The year 2008 has entire year Data

#### Year 2009
```{r y_2009}
summary(yr_2009)
str(yr_2009)
head(yr_2009)
tail(yr_2009)
```
The year 2009 has entire year Data
#### Year 2010
```{r y_2010}
summary(yr_2010)
str(yr_2010)
head(yr_2010)
tail(yr_2010)
```
The year 2010 doesn't have the entire year Data. It has 15 days of missing data.

## Process the Data
create a primary data frame that ONLY includes the data frames that span an entire year.
```{r Combine_tables}
newDF <- bind_rows(yr_2007, yr_2008, yr_2009)
```
yr_2006 and yr_2010 has been excluded, no data for complete year

#### Initial Exploretion of Data set

```{r Initial_Exploretion }
summary(newDF)
str(newDF)
head(newDF,2)
tail(newDF,2)
dim(newDF)
```

#### combine Date and Time to DateTime
```{r new_attribute }
newDF<-mutate(newDF, DateTime = paste(newDF$Date,newDF$Time))
head(newDF,2)
```
 Combine Date and Time attribute values in a new attribute column
#### Move DateTime attribute 
```{r move_attribute }
newDF <- newDF[,c(ncol(newDF), 1:(ncol(newDF)-1))]
head(newDF,2)
```
The column has been moved to first place in the data fram

## remove redundant  features
```{r remove_redundant_att }
newDF <- newDF[,-c(2,3,4,6)]
head(newDF,2)
#newDF$id<-NULL    --another way to remove the feature
```

## rename submeter
```{r rename}
newDF<-newDF%>% dplyr::rename(Kitchen_stuff=Sub_metering_1, Laundry=Sub_metering_2, w_heater_AC=Sub_metering_3) 
head(newDF,2)
#newDF <- rename(newDF, Kitchen_stuff = Sub_metering_1)
```
#### Convert DateTime from POSIXlt to POSIXct 
Convert data type of new DateTime feature
```{r convert}
newDF$DateTime <- as.POSIXct(newDF$DateTime, "%Y/%m/%d %H:%M:%S")
#-Check class of new DateTime feature
class(newDF$DateTime)
tz(newDF$DateTime)
head(newDF,2)

## Add the time zone
attr(newDF$DateTime, "tzone") <- "Europe/Paris"
class(newDF$DateTime)
tz(newDF$DateTime)
head(newDF,2)
tail(newDF,2)
#-check range of time covered by data set
range(newDF$DateTime)
```
#### Assess missing values
```{r missing_values}
sum(is.na(newDF))
#aggr_plot <- aggr(newDF, col=c('navyblue','red'), 
                  #numbers=TRUE, 
                  #sortVars=TRUE, 
                  #labels=names(newDF),
                  #cex.axis=.7,
                  #gap=3, 
                  #ylab=c("Histogram of missing data","Pattern"), digits=2)
#newDF_NA <- newDF[rowSums(is.na(newDF))>0,]
#newDF <- na.omit(newDF)
#sum(is.na(newDF))
```
##### Add feature representing remaining active energy consumed every minute (watt hour)
```{r remaining_energy}
newDF <- newDF %>%
         mutate(Engy_remain=(Global_active_power*1000/60)-
           `Kitchen_stuff` - `Laundry` - `w_heater_AC`)
head(newDF,2)
```
##### Create tidy tibble
In the data set’s current form, each observation or row contains data for each of the sub-meters. To aid the visualization of the submeter data on the same chart, the gather() function is used to create a new column ‘Meter’ which contains the names of the submeters. The observed values for the submeters are placed in the newly-created ‘Watt_hr’ column
```{r tidy}
newDF_tidy <- newDF %>%
  gather(Meter, Watt_hr,  `Kitchen_stuff`,`Laundry`,`w_heater_AC`,`Engy_remain`)
#newDF_tidy %>% as_tibble(newDF_tidy)
               #is_tibble(newDF_tidy)

newDF_tidy$Meter <- factor(newDF_tidy$Meter) #change data type character to factor
head(newDF_tidy %>% as_tibble(newDF_tidy),2)
#head(glimpse(newDF_tidy),2)
```
##  Explore the Data

#### Visualizations of Energy Usage Across Sub-Meters and Time Periods
###### Year_Proportional Plot
```{r Year_Proportional}
Year_Proportional<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(year(DateTime), Meter) %>% 
  summarise(sum=sum(Watt_hr)) %>% 
  ggplot(aes(x=factor(`year(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Year', y='Proportion of Energy Usage') +
  ggtitle('Proportion of Total Yearly Energy Consumption') +
  geom_bar(stat='identity',  position='fill', color='black') 
  #theme(panel.border=element_rect(colour='black', fill=NA)) +
  #theme(text = element_text(size = 14))
Year_Proportional
```

###### Year Bar plot
```{r Year_bar}
Year_bar<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(year(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr/1000),3)) %>%
  ggplot(aes(x=factor(`year(DateTime)`), y=sum)) +
  labs(x='Year', y='kWh') +
  ggtitle('Total Energy Useage by Year') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black')
Year_bar
```
###### Year_Line Plot
```{r Year_line}
Year_line<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(year(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(`year(DateTime)`), sum, group=Meter, colour=Meter)) +
  labs(x='Year', y='kWh') +
  ggtitle('Total Yearly Energy Consumption') +
  geom_line(size=1)
Year_line
```
###### Quarter bar  plot
```{r Quarter_bar}
Quarter_bar<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>% #some missing value in 2010
  #mutate(quarter=quarter(DateTime)) %>%
  group_by(quarter(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr/1000),3)) %>%
  ggplot(aes(x=factor(`quarter(DateTime)`), y=sum)) +
  labs(x='Quarter of the Year', y='kWh') +
  ggtitle('Total Quarterly Energy Consumption') +
  geom_bar(position="dodge",stat='identity', aes(fill = Meter), color='black')
Quarter_bar
```

###### Quarter Proportion plot
```{r Quarter_Proportion}
Quarter_Proportion<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(quarter(DateTime), Meter) %>%
  #filter(quarter(DateTime)<3) %>%
  summarise(sum=round(sum(Watt_hr)/1000),3, na.rm=TRUE) %>%
  ggplot(aes(x=factor(`quarter(DateTime)`), sum, group = Meter, fill=Meter)) +
  labs(x='Quarter of the Year', y='kWh') +
  ggtitle('Proportion of Total Quarterly Energy Consumption') +
  geom_bar(stat='identity', position='fill', color='black')
Quarter_Proportion
```

###### Month Proportional Plot
```{r Month_Proportional}
Month_Proportional<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  mutate(Month=lubridate::month(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Month, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Month), sum, group=Meter,fill=Meter)) +
  labs(x='Month of the Year', y='Proportion of Monthly Energy Useage') +
  ggtitle('Proportion of Total Monthly Energy Useage') +
  geom_bar(stat='identity', position='fill', color='black')
Month_Proportional
```

##### Month Line Plot
```{r Month_Line }
  Month_line<-newDF_tidy %>%
  mutate(Month=lubridate::month(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Month, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Month), sum, group=Meter, colour=Meter)) +
  labs(x='Month of the Year', y='kWh') +
  ggtitle('Average Monthly Energy Useage') +
  geom_line(size=1) +
  geom_line()
Month_line
```

##### Month bar chart
```{r Month_bar }
  Month_bar<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  mutate(Month=lubridate::month(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Month, Meter) %>%
  summarise(sum=round(sum(Watt_hr)/1000),3) %>%
  ggplot(aes(x=factor(Month), y=sum)) +
  labs(x='Month of the Year', y='kWh') +
  ggtitle('Total Energy Useage by Month of the Year') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black')
Month_bar
```
##### Week of the year- line plot
```{r week_line }
  week_line<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(week(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr)/1000),3) %>%
  ggplot(aes(x=factor(`week(DateTime)`), sum, group=Meter,colour=Meter)) +
  labs(x='Week of the Year', y='kWh') +
  ggtitle('Total Energy Usage by Week of the Year') +
  theme(axis.text.x = element_text(angle=90)) +
  geom_line(size=1) +
  geom_line()
week_line
```
##### Week of the year- bar plot
```{r week_bar }
week_bar<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(week(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(`week(DateTime)`), y=sum)) +
  labs(x='Week of the Year', y='kWh') +
  ggtitle('Total Energy Usage by Week of the Year') +
  theme(axis.text.x = element_text(angle=90)) +
  geom_bar(stat='identity', aes(fill=Meter), colour='black')
week_bar
```
##### Day of the Month- Line Plot
#Not informative
```{r week_line_m }
day_line<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(day(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(`day(DateTime)`), sum, group=Meter,colour=Meter)) +
  labs(x='Day of the Month', y='Avg Watt Hour Useage') +
  ggtitle('Average Daily Watt Hour Useage') +
  geom_line(size=1) +
  geom_line()
day_line
```
##### Day of Week- Porportional Plot
```{r week_line_d }
D_W_portion<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  mutate(Day=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Day, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Day), sum, group=Meter,fill=Meter)) +
  labs(x='Day of the Week', y='Proportion of Energy Useage') +
  ggtitle('Proportion of Total Daily Energy Consumption') +
  geom_bar(stat='identity', position='fill', color='black')
D_W_portion
```
##### Day of Week- line
```{r D_W_line }
D_W_line<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  mutate(Day=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Day, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Day), sum, group=Meter, colour=Meter)) +
  labs(x='Day of the Week', y='kWh') +
  ggtitle('Average Daily Energy Consumption') +
  geom_line(size=1) +
  geom_line()
D_W_line
```
##### Day of Week bar chart
```{r week_line_d }
D_W_bar<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  mutate(Day=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Day, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Day), y=sum)) +
  labs(x='Day of the Week', y='kWh') +
  ggtitle('Total Energy Useage by Day of Week') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black')
D_W_bar
```

##### Weekend bar chart
```{r week_line_d }
Weekend_bar <-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  mutate(weekend=lubridate::wday(DateTime, label=TRUE, abbr=FALSE)) %>%
  filter(weekend==c('Saturday','Sunday')) %>%
  group_by(weekend, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(weekend), y=sum, group=Meter,fill=Meter)) +
  labs(x='Weekend Day', y='Proportion of Energy Useage') +
  ggtitle('Proportion of Average Energy Consumption by Weekend Day') +
  geom_bar(stat='identity', position='fill', color='black')
Weekend_bar

```
##### Hour of the Day- Proportional Plot
```{r week_line_d }
Hour_day<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(hour(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(`hour(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Hour of the Day', y='Proportion of Energy Useage') +
  ggtitle('Proportion of Average Hourly Energy Consumption') +
  geom_bar(stat='identity', position='fill', color='black')
Hour_day

```
##### Hour of the Day-Line Plot
```{r week_line_d }
Hour_day_line<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(hour(DateTime), Meter) %>%
  summarise(sum=round(mean(Watt_hr), 3)) %>%
  ggplot(aes(x=factor(`hour(DateTime)`), sum, group=Meter,colour=Meter)) +
  labs(x='Hour of the Day', y='Wh') +
  ggtitle('Average Hourly Energy Consumption') +
  geom_line(size=1) +
  geom_line()
Hour_day_line

```

##### Hour of day bar chart
```{r week_line_d }
Hour_day_bar<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  group_by(hour(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr)/1000),3) %>%
  ggplot(aes(x=factor(`hour(DateTime)`), y=sum)) +
  labs(x='Hour of the Day', y='kWh') +
  ggtitle('Total Energy Useage by Hour of the Day') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black')
Hour_day_bar

```
#### Winter
###### Filter and plot data for weeks 1-8
```{r week_line_d }
winter_plot<-newDF_tidy %>%
  filter(year(DateTime)<2010) %>%
  filter(week(DateTime) == c(1:8)) %>%
  mutate(Day=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Day, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Day), y=sum)) +
  labs(x='Day of the Week', y='kWh') +
  ylim(0,200) +
  ggtitle('Total Energy Usage by Day for Weeks of \nHigh Consumption in Winter Months') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black') +
  theme(panel.border=element_rect(colour='black', fill=NA)) +
  theme(text = element_text(size = 14))
winter_plot
```

#### Summer
###### Filter and plot data for weeks 18-25
```{r week_line_d }
summer_plot<-newDF_tidy %>%
  filter(week(DateTime) == c(18:25)) %>%
  mutate(Day=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Day, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Day), y=sum)) +
  labs(x='Day of the Week', y='kWh') +
  #ylim(0,85) +
  ggtitle('Total Energy Usage by Day for Weeks of \nHigh Consumptionin Summer Months') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black') +
  theme(panel.border=element_rect(colour='black', fill=NA)) +
  theme(text = element_text(size = 14))
summer_plot
#sum(is.na(newDF_tidy$Watt_hr))
```
### Summary Plot(Summer and Winter)