---
title: "Analysis of Submetered Household Electric Power Consumption"
author: "Ashiqur"
date: "11/20/2019"
output: html_document
---
```{r setup, include=TRUE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

```{r library}
library(RMySQL)
library(dplyr)
library(lubridate)
library(VIM)        #Aids visualization and imputing of missing values
library(funModeling)
library(ggplot2)
library(reshape2)
library(caret)      #R modeling workhorse & ggplot2
library(tidyverse)  #Package for tidying datalibrary(magrittr)   #Enables piping
library(GGally)
```

## Data Connection & Retrieving Data
##### Create a database connection 
```{r load_data}
con = dbConnect(MySQL(), user='deepAnalytics', 
                password='Sqltask1234!', 
                dbname='dataanalytics2018', 
                host='data-analytics-2018.cbrosir2cswx.us-east-1.rds.amazonaws.com')
```

### List the tables contained in the database
```{r table}
dbListTables(con)
```
### Lists attributes contained in a table
```{r attributes}
dbListFields(con,'yr_2006')
```

## Request relevant data base tables

#### Use asterisk to specify all attributes for download

```{r pressure}
yr_2006 <- dbGetQuery(con, "SELECT * FROM yr_2006")
yr_2007 <- dbGetQuery(con, "SELECT * FROM yr_2007")
yr_2008 <- dbGetQuery(con, "SELECT * FROM yr_2008")
yr_2009 <- dbGetQuery(con, "SELECT * FROM yr_2009")
yr_2010 <- dbGetQuery(con, "SELECT * FROM yr_2010")
```
## Investigate each new data frame

#### Year 2006
```{r y_2006}
summary(yr_2006)
str(yr_2006)
head(yr_2006)
tail(yr_2006)
```
The year 2006 doesn't have the entire year Data. It has only 15 days of data.

#### Year 2007
```{r y_2007}
summary(yr_2007)
str(yr_2007)
head(yr_2007)
tail(yr_2007)
```
The year 2007 has entire year Data

#### Year 2008
```{r y_2008}
summary(yr_2008)
str(yr_2008)
head(yr_2008)
tail(yr_2008)
```
The year 2008 has entire year Data

#### Year 2009
```{r y_2009}
summary(yr_2009)
str(yr_2009)
head(yr_2009)
tail(yr_2009)
```
The year 2009 has entire year Data
#### Year 2010
```{r y_2010}
summary(yr_2010)
str(yr_2010)
head(yr_2010)
tail(yr_2010)
```
The year 2010 doesn't have the entire year Data. It has 15 days of missing data.

## Data Preparation
create a primary data frame that ONLY includes the data frames that span an entire year.
```{r Combine_tables}
newDF <- bind_rows(yr_2007, yr_2008, yr_2009)
```
yr_2006 and yr_2010 has been excluded, no data for complete year

#### Initial Exploretion of Data set

```{r Initial_Exploretion }
summary(newDF)
str(newDF)
head(newDF,2)
tail(newDF,2)
dim(newDF)
```

#### combine Date and Time to DateTime
```{r new_attribute }
newDF<-mutate(newDF, DateTime = paste(newDF$Date,newDF$Time))
head(newDF,2)
```
 Combine Date and Time attribute values in a new attribute column
#### Move DateTime attribute 
```{r move_attribute }
newDF <- newDF[,c(ncol(newDF), 1:(ncol(newDF)-1))]
head(newDF,2)
```
The column has been moved to first place in the data fram

## remove redundant  features
```{r remove_redundant_att }
newDF <- newDF[,-c(2,3,4,6)]
head(newDF,2)
#newDF$id<-NULL    --another way to remove the feature
```

## rename submeter
```{r rename}
newDF<-newDF%>% dplyr::rename(Kitchen_stuff=Sub_metering_1, Laundry=Sub_metering_2, w_heater_AC=Sub_metering_3) 
head(newDF,2)
#newDF <- rename(newDF, Kitchen_stuff = Sub_metering_1)
```
#### Convert DateTime from POSIXlt to POSIXct 
Convert data type of new DateTime feature
```{r convert}
newDF$DateTime <- as.POSIXct(newDF$DateTime, "%Y/%m/%d %H:%M:%S")
#-Check class of new DateTime feature
class(newDF$DateTime)
tz(newDF$DateTime)
head(newDF,2)

## Add the time zone----------
attr(newDF$DateTime, "tzone") <- "Europe/Paris"
class(newDF$DateTime)
tz(newDF$DateTime)
head(newDF,2)
tail(newDF,2)
```