---
title: "Time_series_Energy_data"
author: "Ashiqur"
date: "11/28/2019"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = TRUE)
```
## Load Packages

```{r library}
library(RMySQL)
library(dplyr)
library(lubridate)
library(VIM)        #Aids visualization and imputing of missing values
library(funModeling)
library(ggplot2)
library(reshape2)
library(caret)      #R modeling workhorse & ggplot2
library(tidyverse)  #Package for tidying datalibrary(magrittr)   #Enables piping
library(GGally)
library(plotly)
library(ggfortify)
library(TTR)        #Decomposing time series
library(forecast)
```
## Data Connection & Retrieving Data
##### Create a database connection 
```{r load_data}
con = dbConnect(MySQL(), user='deepAnalytics', 
                password='Sqltask1234!', 
                dbname='dataanalytics2018', 
                host='data-analytics-2018.cbrosir2cswx.us-east-1.rds.amazonaws.com')
```

### List the tables contained in the database
```{r table}
dbListTables(con)
```
### Lists attributes contained in a table
```{r attributes}
dbListFields(con,'yr_2006')
```

## Request relevant data base tables

#### Use asterisk to specify all attributes for download

```{r pressure}
yr_2006 <- dbGetQuery(con, "SELECT * FROM yr_2006")
yr_2007 <- dbGetQuery(con, "SELECT * FROM yr_2007")
yr_2008 <- dbGetQuery(con, "SELECT * FROM yr_2008")
yr_2009 <- dbGetQuery(con, "SELECT * FROM yr_2009")
yr_2010 <- dbGetQuery(con, "SELECT * FROM yr_2010")
```
## Load Data
create a primary data frame that ONLY includes the data frames that span an entire year.
```{r Combine_tables}
newDF <- bind_rows(yr_2007, yr_2008, yr_2009)
```
yr_2006 and yr_2010 has been excluded, no data for complete year

## rename submeter
```{r rename}
newDF<-newDF%>% dplyr::rename(Kitchen_stuff=Sub_metering_1, Laundry=Sub_metering_2, w_heater_AC=Sub_metering_3) 
head(newDF,2)
#newDF <- rename(newDF, Kitchen_stuff = Sub_metering_1)
```

## visualization 
One of the goals of subsetting for visualizations is to adjust granularity to maximize the information to be gained. Granularity describes the frequency of observations within a time series data set

#### Convert DateTime from POSIXlt to POSIXct 
Convert data type of new DateTime feature
```{r convert}
newDF<-mutate(newDF, DateTime = paste(newDF$Date,newDF$Time))
newDF$DateTime <- as.POSIXct(newDF$DateTime, "%Y/%m/%d %H:%M:%S")
#-Check class of new DateTime feature
class(newDF$DateTime)
tz(newDF$DateTime)
head(newDF,2)

## Add the time zone
attr(newDF$DateTime, "tzone") <- "Europe/Paris"
class(newDF$DateTime)
tz(newDF$DateTime)
head(newDF,2)
tail(newDF,2)
#-check range of time covered by data set
range(newDF$DateTime)
```

```{r Granularity}
plot(newDF$Kitchen_stuff)
head(newDF,2)
```
Over 1,500,000 minutes from the beginning of 2007 though the end of 2009 have been plotted in this visualization.we are not getting any usefull information.

##### Subsetting and Meaningful Time Periods
```{r subset_week}
## Subset the second week of 2008 - All Observations
houseWeek <- filter(newDF, year(DateTime) == 2008 & week(DateTime) == 2)
## Plot subset houseWeek
plot(houseWeek$Kitchen_stuff)
```
There are lots of patterns to be learned from a week, but the granularity has to be right to extract this information. The ‘week’ plot above has over 10,000 observations and this granularity should probably be reduced

#### Visualize a Single Day with Plotly
```{r subset_day}
## Subset the second week of 2008 - All Observations
houseDay <- filter(newDF, year(DateTime) == 2008 & week(DateTime) == 2 & day(DateTime)==9)
## Plot kitchen_stuff
plot(houseDay$Kitchen_stuff)
plot_ly(houseDay, x = ~houseDay$DateTime, y = ~houseDay$Kitchen_stuff, type = 'scatter', mode = 'lines')
```
line plots and Plotly make much better visualizations than the basic plot function.

#### 3 submeter together
Perhaps the best way to understand the power usage on this day is to plot all three sub-meters to gain a perspective on all of the power being consumed. While we're at it let's add a legend, title and axis labels.  
```{r subset_day_3}
plot_ly(houseDay, x = ~houseDay$DateTime, y = ~houseDay$Kitchen_stuff, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseDay$Laundry, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseDay$w_heater_AC, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January 9th, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```

#### Reducing Granularity
```{r subset_day_3_10}
houseDay_10 <- filter(newDF, year(DateTime) == 2008 & week(DateTime) == 2 & day(DateTime)==9 & (minute(DateTime) == 0 | minute(DateTime) == 10 | minute(DateTime) == 20 | minute(DateTime) == 30 | minute(DateTime) == 40 | minute(DateTime) == 50))
plot_ly(houseDay_10, x = ~houseDay_10$DateTime, y = ~houseDay_10$Kitchen_stuff, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseDay_10$Laundry, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseDay_10$w_heater_AC, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January 9th, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```
With the granularity adjusted we get a much more clear picture of the power consumption on January 9th

#### week of Day
```{r Day}
#mutate(Day=lubridate::wday(DateTime, label=TRUE, abbr=TRUE))


houseDay_W <- filter(newDF, year(DateTime) == 2008 & month(DateTime)== 1 & week(DateTime) == 2 & wday(DateTime))

plot_ly(houseDay_W, x = ~houseDay_W$DateTime, y = ~houseDay_W$Kitchen_stuff, name = 'Kitchen', type = 'scatter', mode = 'lines') %>%
 add_trace(y = ~houseDay_W$Laundry, name = 'Laundry Room', mode = 'lines') %>%
 add_trace(y = ~houseDay_W$w_heater_AC, name = 'Water Heater & AC', mode = 'lines') %>%
 layout(title = "Power Consumption January 9th, 2008",
 xaxis = list(title = "Time"),
 yaxis = list (title = "Power (watt-hours)"))
```

## Convert to Time Series and Plot

##### Subset to one observation per week on Mondays at 8:00pm for 2007, 2008 and 2009

```{r house070809weekly}
house070809weekly <-
        filter(newDF,
               weekdays(DateTime) == "Monday" &
                       hour(DateTime) == 20 & minute(DateTime) == 1)

## Create TS object with SubMeter3
tsSM3_070809weekly <- ts(house070809weekly$w_heater_AC, 
                         frequency=52, 
                         start=c(2007,1),
                         end=c(2009,52))
autoplot(tsSM3_070809weekly, ts.colour = 'red', xlab = "Time", ylab = "Watt Hours", main = "Sub-meter 3")
#autoplot(tsSM3_070809weekly)
## Plot sub-meter 3 with plot.ts
plot.ts(tsSM3_070809weekly)

```

#### Create monthly time series
```{r house070809weekly_1}
house_monthly <- filter(newDF,year(DateTime)==2009 & month(DateTime)==1) %>% 
        group_by(day(DateTime)) %>% 
        summarise(kitchen = sum(Kitchen_stuff))

## Create TS object with SubMeter3
tsSM1_house_monthly <- ts(house_monthly$kitchen, 
                         frequency=365, 
                         start=c(2009,1),
                         end=c(2009,31))

plot.ts(tsSM1_house_monthly, xlab = "Time", ylab = "Watt Hours", main = "Kitchen_stuff")
#autoplot(tsSM3_070809weekly)
## Plot sub-meter 3 with plot.ts
#plot.ts(tsSM1_house_monthly)

```

##### Subset data by year and summarise total usage across the 3 submeters(Time series)
```{r housePWR_yr}
  housePWR_yr <- newDF %>%
  filter(year(DateTime)<2010) %>%
  group_by(year(DateTime)) %>%
  summarise(Kitchen_stuff=round(sum(`Kitchen_stuff`/1000, na.rm=TRUE), 3),
            Laundry=round(sum(`Laundry`/1000, na.rm=TRUE), 3),
            w_heater_AC=round(sum(`w_heater_AC`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime))
housePWR_yr
housePWR_yrTS <- ts(housePWR_yr[,2:4], 
                    frequency = 1, 
                    start=c(2007))
plot(housePWR_yrTS, plot.type='s', #xaxt='n',
     #xaxp = c(2007, 2010, 3),
     col=c('red', 'green', 'blue'),
     main='Total Yearly kWh Consumption (2007-2010)',
     xlab='Year', ylab = 'kWh')
b <- c('Kitchen_stuff', 'Laundry', 'w_heater_AC')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')
```

##### Subset data by Semester and summarise total usage across the 3 submeters(Time series)
```{r housePWR_semstr}
  housePWR_semstr <- newDF %>%
  filter(year(DateTime)<2010) %>%
  group_by(year(DateTime),semester(DateTime)) %>%
  summarise(Kitchen_stuff=round(sum(`Kitchen_stuff`/1000, na.rm=TRUE), 3),
            Laundry=round(sum(`Laundry`/1000, na.rm=TRUE), 3),
            w_heater_AC=round(sum(`w_heater_AC`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime))
housePWR_semstr

housePWR_semstrTS <- ts(housePWR_semstr[,3:5], 
                        frequency = 2, 
                        start=c(2007,1),
                        end=c(2010))
plot(housePWR_semstrTS, plot.type='s', #xaxt='n',
     #xaxp = c(2007, 2010, 3),
     col=c('red', 'green', 'blue'),
     main='Total Energy Consumption by Semester',
     xlab='Year', ylab = 'kWh')
b <- c('Kitchen_stuff', 'Laundry', 'w_heater_AC')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')

```

##### Subset data by quarter and summarise total usage across the 3 submeters(Time series)
```{r housePWR_qtr}
housePWR_qtr <- newDF %>%
          filter(year(DateTime)<2010) %>%
  group_by(year(DateTime), quarter(DateTime)) %>%
  summarise(Kitchen_stuff=round(sum(`Kitchen_stuff`/1000), 3),
            Laundry =round(sum(`Laundry`/1000), 3),
            w_heater_AC=round(sum(`w_heater_AC`/1000), 3))
housePWR_qtr
housePWR_qtr_sma<- newDF %>%
          filter(year(DateTime)<2010) %>%
  group_by(year(DateTime), quarter(DateTime)) %>%
  summarise(w_heater_AC=round(sum(`w_heater_AC`/1000), 3))

housePWR_qtrTS <- ts(housePWR_qtr[,3:5],
                     frequency=4,
                     start=c(2007,1),
                     end=c(2010))
plot(housePWR_qtrTS, 
     plot.type='s',
     col=c('red', 'green', 'blue'),
     main='Total Quarterly kWh Consumption (2007-2010)',
     xlab='Year', ylab = 'kWh')
minor.tick(nx=4)
#-Create legend
b <- c('Kitchen_stuff', 'Laundry', ' w_heater_AC')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')
```

##### Subset data by month and summarise total usage across the 3 submeters(Time series)
```{r housePWR_mnth}
housePWR_mnth <- newDF %>%
  group_by(year(DateTime), month(DateTime)) %>%
  summarise(Kitchen_stuff=round(sum(`Kitchen_stuff`/1000), 3),
            Laundry=round(sum(`Laundry`/1000), 3),
            w_heater_AC=round(sum(`w_heater_AC`/1000), 3))
head(housePWR_mnth)
housePWR_mnthTS <- ts(housePWR_mnth[,3:5],
                      frequency = 12,
                      start=c(2007,1),
                      end=c(2009,12))
plot(housePWR_mnthTS, 
     plot.type='s',
     xlim=c(2007, 2010),
     col=c('red', 'green', 'blue'),
     main='Total Monthly kWh Consumption',
     xlab='Year/Month', ylab = 'kWh')
minor.tick(nx=12)
b <- c('Kitchen_stuff', 'Laundry', 'w_heater_AC')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')
```

##### Subset data by week of year and summarise total usage across the 3 submeters(Time series)
```{r housePWR_wkofYr}
housePWR_wkofYr <- newDF %>%
  #filter(year(DateTime)>2006) %>%
  filter(year(DateTime)<2010) %>%
  mutate(DofWk=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(year(DateTime),week(DateTime)) %>%
  summarise(Kitchen_stuff=round(sum(`Kitchen_stuff`/1000, na.rm=TRUE), 3),
            Laundry=round(sum(`Laundry`/1000, na.rm=TRUE), 3),
            w_heater_AC=round(sum(`w_heater_AC`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime))
housePWR_wkofYr
housePWR_wkofYrTS <- ts(housePWR_wkofYr[,3:5], 
                        frequency =53, start=c(2007,1), 
                        end=c(2010))
plot(housePWR_wkofYrTS, plot.type='s', #xaxt='n',
     #xaxp = c(1, 13, 12),
     col=c('red', 'green', 'blue'),
     xlab ='Year', ylab = 'kWh',
     #ylim=c(0,120),
     main='Total Energy Consumption by Week of the Year')
#axis(side=1, at= c(1, 2,3,4,5,6,7,8,9,10,11,12,13), labels=MonthLst)
minor.tick(nx=52)
b <- c('Kitchen_stuff', 'Laundry', 'w_heater_AC')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')


```

##### Subset data by day of week and summarise total usage across the 3 submeters(Time series)
```{r housePWR_dofWk }
housePWR_dofWk <- newDF %>%
  #filter(year(DateTime)>2006) %>%
  filter(year(DateTime)<2010) %>%          
  mutate(DofWk=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(week(DateTime),DofWk) %>%
  #group_by(lubridate::week(DateTime), lubridate::year(DateTime)) %>%
  summarise(Kitchen_stuff=round(sum(`Kitchen_stuff`/1000, na.rm=TRUE), 3),
            Laundry=round(sum(`Laundry`/1000, na.rm=TRUE), 3),
            w_heater_AC=round(sum(`w_heater_AC`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime),
            last_DateTime = last(DateTime)) %>% 
        arrange(first_DateTime)
head(housePWR_dofWk)
#tail(housePWR_dofWk)
#Day of Week_TS
housePWR_dofWkTS <- ts(housePWR_dofWk[,3:5], 
                       frequency =7, 
                       start=c(1,1))
plot(housePWR_dofWkTS, plot.type='s', #xaxt='n',
     col=c('red', 'green', 'blue'),
     xlab ='Week of Year', ylab = 'kWh',
     xlim = c(1,53) , ylim=c(0,75),
     main='Total Energy Consumption by Day of Week')
minor.tick(nx=7)
b <- c('Kitchen_stuff', 'Laundry', 'w_heater_AC')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')


```

##### Subset data by hour of day and summarise total usage across the 3 submeters(Time series)
```{r housePWR_hofDay }
housePWR_hofDay <- newDF %>%

  #filter((minute(DateTime) %% 5) == 0) %>%
  group_by(wday(DateTime), hour(DateTime)) %>%
  summarise(Kitchen_stuff=round(sum(`Kitchen_stuff`/1000, na.rm=TRUE), 3),
            Laundry=round(sum(`Laundry`/1000, na.rm=TRUE), 3),
            w_heater_AC=round(sum(`w_heater_AC`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime))
housePWR_hofDay
housePWR_hofDayTS <- ts(housePWR_hofDay[,3:5], 
                        frequency=24, 
                        start=c(0,0))
plot(housePWR_hofDayTS, plot.type='s',
     #xaxp = c(0,48, 47),
     col=c('red', 'green', 'blue'),
     xlab='Day', ylab = 'kWh',
     main='Total kWh Consumption by Hour of the Day')
minor.tick(nx=24)
b <- c('Kitchen_stuff', 'Laundry', 'w_heater_AC')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')

```

##### Subset data by weekend of day and summarise total usage across the 3 submeters(Time series)
```{r }
housePWR_wknd <- newDF %>%
  #filter(year(DateTime)>2006) %>%
  #filter(year(DateTime)<2010) %>%
  mutate(Wknd=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  filter(Wknd == c('Sat', 'Sun')) %>%
  group_by(Wknd, hour(DateTime)) %>%
  summarise(Kitchen_stuff=round(sum(`Kitchen_stuff`/1000, na.rm=TRUE), 3),
            Laundry=round(sum(`Laundry`/1000, na.rm=TRUE), 3),
            w_heater_AC=round(sum(`w_heater_AC`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime)) %>%
  arrange(desc(Wknd))
housePWR_wknd

housePWR_wkndTS <- ts(housePWR_wknd[,3:5], frequency=24, end(1,23))
class(housePWR_wkndTS)
plot(housePWR_wkndTS, plot.type='s', xaxt='n',
    # xaxp = c(0, 3,2),
     xlim=c(1,3),
     col=c('red', 'green', 'blue'),
     xlab='Weekend Day', ylab = 'kWh',
     ylim=c(0,90),
     main='Total Weekend Energy Consumption by Hour')
#axis(side = 1, at = c(1,2,3), labels = WkndList)
minor.tick(nx=24)
b <- c('Kitchen_stuff', 'Laundry', 'w_heater_AC')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')
```
## Fit Moving Average Model to Quarterly Time Series
```{r Moving_Average}
# plotting time series
plot.ts(housePWR_qtrTS)
# Transform of time series in case of seasonal and  random flactuation 
housePWR_qtrTS_log<-log(housePWR_qtrTS)
dim(housePWR_qtrTS_log)
plot.ts(housePWR_qtrTS_log)
# In case fluctuations are present  and you  want to smooth the time series

#housePWR_qtrTS<-housePWR_qtrTS[,-c(1:2)]
#housePWR_qtrTS

housePWR_qtrTS_sma3<-SMA(housePWR_qtrTS_log[,3],3)
plot.ts(housePWR_qtrTS_sma3)


housePWR_qtrTS_sma5<-SMA(housePWR_qtrTS_log[,3],5)
plot.ts(housePWR_qtrTS_sma5)
# it's giving a some good results and showing there is a trand present.so its ok for me to do the further analysis prediction based on that 

housePWR_qtrTS_sma7<-SMA(housePWR_qtrTS_log[,3],7)
plot.ts(housePWR_qtrTS_sma7)
#more period i am using more data point i will going to lose


```
## Exponential Smoothing
##### Fit Exponential Moving Average Model to Quarterly Time Series
```{r Exponential}
# plotting time series
plot.ts(housePWR_qtrTS)
# Transform of time series in case of seasonal and  random flactuation 
housePWR_qtrTS_log<-log(housePWR_qtrTS)
dim(housePWR_qtrTS_log)
plot.ts(housePWR_qtrTS_log)
# In case fluctuations are present  and you  want to smooth the time series

#housePWR_qtrTS<-housePWR_qtrTS[,-c(1:2)]
#housePWR_qtrTS

housePWR_qtrTS_ema_25<-EMA(housePWR_qtrTS_log[,3],3,ratio = .25)
plot.ts(housePWR_qtrTS_ema_25)

housePWR_qtrTS_ema_5<-EMA(housePWR_qtrTS_log[,3],3,ratio = .5)
plot.ts(housePWR_qtrTS_ema_5)
# it's giving a some good results and showing there is a trand present.so its ok for me to do the further analysis prediction based on that 

housePWR_qtrTS_ema_75<-EMA(housePWR_qtrTS_log[,3],3,ratio = .75)
plot.ts(housePWR_qtrTS_ema_75)
#more period i am using more data point i will going to lose
```
In this data set simple moving average is kind of a winner for me because with that i am able to see very clearly there is a trand that is present,there is some cyclic or random or seasonal that is present but with this Exponential Moving Average i could only see that probly there is a trand but there is an irregular component which is present in this time series.

## Decomposition
Decomposing the time series means separating the time series into these three components:
```{r Decomposition}
# plotting time series
plot.ts(housePWR_qtrTS)
housePWR_qtrTS_decomp<-decompose(housePWR_qtrTS[,3])
plot(housePWR_qtrTS_decomp)
summary(housePWR_qtrTS_decomp)

#if random component want to remove 

housePWR_qtrTS_random <- (housePWR_qtrTS[,3]) - housePWR_qtrTS_decomp$random
plot(housePWR_qtrTS_random)
```

# Holt-Winter Forecasting
To make forecasts using simple exponential smoothing, you can fit a simple exponential smoothing predictive model using the HoltWinters() function from the stats package for R. 
```{r remove_seasonal}
#plot.ts(housePWR_qtrTS)
#Remove Seasonal Components
housePWR_qtrTS_seasonal <- (housePWR_qtrTS[,3]) - housePWR_qtrTS_decomp$seasonal
plot.ts(housePWR_qtrTS_seasonal)
#autoplot(housePWR_qtrTS_seasonal)
## Test Seasonal Adjustment by running Decompose again. Note the very, very small scale for Seasonal
plot(decompose(housePWR_qtrTS_seasonal))
```

Yes there is a seasonal line, but look at the scale for the seasonal section. -1e-15 through 5e-16. That's a decimal with 15 zeros before 1. A very very small number indeed. For all practical purposes the seasonality has been removed. 

##### HoltWinters Simple Exponential Smoothing
```{r HW_smoothing}
## Holt Winters Exponential Smoothing & Plot
plot.ts(housePWR_qtrTS[,3])
housePWR_qtrTS_HW_smoothing<-HoltWinters(housePWR_qtrTS[,3], beta=FALSE, gamma=FALSE)
housePWR_qtrTS_HW_smoothing

checkresiduals(housePWR_qtrTS_HW_smoothing)

housePWR_qtrTS_HW_smoothing$fitted
plot(housePWR_qtrTS_HW_smoothing)

## HoltWinters forecast & plot
housePWR_qtrTS_HW_smoothing_forec <- forecast(housePWR_qtrTS_HW_smoothing, h=4)
plot(housePWR_qtrTS_HW_smoothing_forec, ylim = c(0, 1300), ylab= "Watt-Hours", xlab="Time - w_heater_ac")

housePWR_qtrTS_HW_smoothing_onlyforec <- forecast(housePWR_qtrTS_HW_smoothing, h=4,level=c(12,16))
plot(housePWR_qtrTS_HW_smoothing_forec, ylim = c(0, 1300), ylab= "Watt-Hours", xlab="Time - w_heater_ac",start(2010))
plot.ts(housePWR_qtrTS_HW_smoothing_forec$residuals)


```

Less weight to the recent value and more weight to the previous value 

## Time series tasting

```{r ACF}
## ACF of residuals of Holt Winters forecast
acf(housePWR_qtrTS_HW_smoothing_forec$residuals, 
    lag.max = 10,
    na.action = na.pass)
#acf(na.exclude(housePWR_qtrTS_HW_smoothing_forec$residuals),lag.max=10)
plot(housePWR_qtrTS_HW_smoothing_forec$residuals)
```

No lag basically touching the minimum and maximum value,so There is no significant correlation between the forecast error and the predicted values

```{r pacf}
pacf(housePWR_qtrTS_HW_smoothing_forec$residuals, 
     lag.max = 10,
     na.action = na.pass,
      plot = FALSE)

pacf(housePWR_qtrTS_HW_smoothing_forec$residuals, 
     lag.max = 10,
     na.action = na.pass)

```
```{r Ljung_box}
#also check the for correlation
Box.test(housePWR_qtrTS_HW_smoothing_forec$residuals, 
         lag = 10, 
         type = "Ljung-Box")
plot.ts(housePWR_qtrTS_HW_smoothing_forec$residuals)
hist(housePWR_qtrTS_HW_smoothing_forec$residuals)
```

p-value = 0.2401 that indicate some little bit correlation is present jot very significant .The histogram shows it's not normally distributed(Value is not aggregating near to the zero )

## ARIMA Model
#### Total Monthly kWh Consumption(water heater)
It is based on the assumption that over a preiod of times the current values are related or correlated with their immediate previous OR n previous values.
```{r ARIMA_w_heater_AC }

df_arima_w_heater_AC<-(housePWR_mnthTS[,3]) 
plot(df_arima_w_heater_AC)
#cycle(df_arima_w_heater_AC)
#Box plot 
boxplot(df_arima_w_heater_AC ~ cycle(df_arima_w_heater_AC))#Box plot across months will give us a sense on seasonal effect
#This will fit in a line
#abline(reg=lm(df_arima ~ time(df_arima)))
#we need to remove unequal variance.we do this using log function
plot(log(df_arima_w_heater_AC))
#Mean has to constant according to time (need to do untill unless mean to make constant)
plot(diff(log(df_arima_w_heater_AC)))

df_fit_arima_mnth_TS_auto<-auto.arima(df_arima_w_heater_AC)
#will calculate everything automatic 

#AR I MA
#p  d q this value we have to get from the graph 
#c(p,d,q)

acf(df_arima_w_heater_AC) # determine the value of q=2
#acf(diff(log(df_arima)))

pacf(df_arima_w_heater_AC)  # determine the value of p=0
df_fit_arima_mnth_TS<-arima(df_arima_w_heater_AC, c(0,0,2))

df_fit_arima_mnth_TS_pred<-predict(df_fit_arima_mnth_TS,n.ahead = 1*12)

ts.plot(df_arima_w_heater_AC,df_fit_arima_mnth_TS_pred$pred)
#pred<-2.718^df_fit_arima_mnth_TS_pred$df_fit_arima_mnth_TS_pred--if use log then
```
Mean value is almost constant but variance is not the same

#### Total Monthly kWh Consumption(kitchen)
```{r ARIMA_Kitchen_stuff }

df_arima_Kitchen_stuff<-(housePWR_mnthTS[,1]) 
plot(df_arima_Kitchen_stuff)
#cycle(df_arima_Kitchen_stuff)
#Box plot 
boxplot(df_arima_Kitchen_stuff ~ cycle(df_arima_Kitchen_stuff))#Box plot across months will give us a sense on seasonal effect
#This will fit in a line
#abline(reg=lm(df_arima_Kitchen_stuff ~ time(df_arima)))
#we need to remove unequal variance.we do this using log function
plot(log(df_arima_Kitchen_stuff))
#Mean has to constant according to time (need to do untill unless mean to make constant)
plot(diff(log(df_arima_Kitchen_stuff)))

df_fit_arima_mnth_TS_auto<-auto.arima(df_arima_w_heater_AC)
#will calculate everything automatic 

#AR I MA
#p  d q this value we have to get from the graph 
#c(p,d,q)

acf(diff(log(df_arima_Kitchen_stuff))) # determine the value of q=2
#acf(diff(log(df_arima)))

pacf(diff(log(df_arima_Kitchen_stuff)))  # determine the value of p=0
df_fit_arima_mnth_kitc<-arima(log(df_arima_Kitchen_stuff), c(1,1,2))


df_fit_arima_mnth_TS_pred<-predict(df_fit_arima_mnth_kitc,n.ahead = 1*12)

df_fit_arima_mnth_TS_pred_no_log<-2.718^df_fit_arima_mnth_TS_pred$pred

ts.plot(df_arima_Kitchen_stuff,2.718^df_fit_arima_mnth_TS_pred$pred,log="y",lty=c(1,3))
#pred<-2.718^df_fit_arima_mnth_TS_pred$df_fit_arima_mnth_TS_pred--if use log then
```

#### Total Monthly kWh Consumption(Laundray)
```{r ARIMA_Laundray }

df_arima_Laundry<-(housePWR_mnthTS[,2]) 
plot(df_arima_Laundry)
#cycle(df_arima_Laundray)
#Box plot 
boxplot(df_arima_Laundry ~ cycle(df_arima_Laundry))#Box plot across months will give us a sense on seasonal effect
#This will fit in a line
#abline(reg=lm(df_arima_Kitchen_stuff ~ time(df_arima)))
#we need to remove unequal variance.we do this using log function
plot(log(df_arima_Laundry))

#Mean has to constant according to time (need to do untill unless mean to make constant)
plot(diff(log(df_arima_Laundry)))

df_fit_arima_mnth_TS_auto<-auto.arima(df_arima_Laundry)
#will calculate everything automatic 

#AR I MA
#p  d q this value we have to get from the graph 
#c(p,d,q)

acf(diff(log(df_arima_Laundry))) # determine the value of q=3
#acf(diff(log(df_arima)))

pacf(diff(log(df_arima_Laundry)))  # determine the value of p=2
df_fit_arima_mnth_laun<-arima(log(df_arima_Laundry), c(2,1,3))


df_fit_arima_mnth_laun_pred<-predict(df_fit_arima_mnth_laun,n.ahead = 1*12)

df_fit_arima_mnth_laun_pred_no_log<-2.718^df_fit_arima_mnth_laun_pred$pred

ts.plot(df_arima_Laundry,2.718^df_fit_arima_mnth_laun_pred$pred,log="y",lty=c(1,3))
#pred<-2.718^df_fit_arima_mnth_TS_pred$df_fit_arima_mnth_TS_pred--if use log then
```

## testing Arima model(water Heater)
```{r testing_arima_wh}
df_arima_7_8<-ts(df_arima_w_heater_AC,frequency = 12,
                 start = c(2007,1),
                 end = c(2008,12))
df_arima_7_8_fit<-arima(df_arima_7_8, c(0,0,2))

df_arima_7_8_fit_pred<-predict(df_arima_7_8_fit,n.ahead = 1*12)#2009 predicted value
df_arima_7_8_fit_pred_p<-(df_arima_7_8_fit_pred$pred)


df_arima_7_8_fit_pred_round<-round(df_arima_7_8_fit_pred_p,digits = 0)
df_arima_7_8_fit_pred_round
w_heater_AC<-tail(df_arima_w_heater_AC,12)
w_heater_AC

```
## testing Arima model(kitchen)
```{r testing_arima_kit}
df_arima_7_8_kit<-ts(df_arima_Kitchen_stuff,frequency = 12,
                 start = c(2007,1),
                 end = c(2008,12))
df_arima_7_8_Arima_kit<-arima(log(df_arima_7_8_kit), c(1,1,2))

df_arima_7_8_fit_pred_kit<-predict(df_arima_7_8_Arima_kit,n.ahead = 1*12)#2009 predicted value
df_arima_7_8_fit_pred_p_kit<-(df_arima_7_8_fit_pred_kit$pred)

df_arima_7_8_fit_pred_p_kit_no_log<-2.718^df_arima_7_8_fit_pred_p_kit
df_arima_7_8_fit_pred_p_kit_no_log_round<-round(df_arima_7_8_fit_pred_p_kit_no_log,digits = 0)
df_arima_7_8_fit_pred_p_kit_no_log_round
kitchen<-tail(df_arima_Kitchen_stuff,12)
kitchen
```
## testing Arima model(Laundry)
```{r testing_arima_Laundry}
df_arima_7_8_laun<-ts(df_arima_Laundry,frequency = 12,
                 start = c(2007,1),
                 end = c(2008,12))
df_arima_fit_7_8_laun<-arima(log(df_arima_7_8_laun), c(2,1,3))

df_arima_7_8_fit_pred_laun<-predict(df_arima_fit_7_8_laun,n.ahead = 1*12)#2009 predicted value
df_arima_7_8_fit_pred_p_laun<-(df_arima_7_8_fit_pred_laun$pred)

df_arima_7_8_fit_pred_p_laun_no_log<-2.718^df_arima_7_8_fit_pred_p_laun
df_arima_7_8_fit_pred_p_laun_no_log_round<-round(df_arima_7_8_fit_pred_p_laun_no_log,digits = 0)
df_arima_7_8_fit_pred_p_laun_no_log_round
Laundra_y<-tail(df_arima_Laundry,12)
Laundra_y
```
##  Forecast ARIMA model(water heater)
```{r Forcast_arima_wh_ac}
#will calculate everything automatic 
df_fit_arima_mnth_TS_auto<-auto.arima(df_arima_w_heater_AC)
Box.test(df_fit_arima_mnth_TS_auto$residuals,
         lag=20,
         type="Ljung-Box")
hist(df_fit_arima_mnth_TS_auto$residuals)
#lines(density(df_fit_arima_mnth_TS_auto$residuals))
checkresiduals(df_fit_arima_mnth_TS_auto)

df_fit_arima_mnth_TS_auto_forecast<-forecast(df_fit_arima_mnth_TS_auto,h=12)
autoplot(df_fit_arima_mnth_TS_auto_forecast)
accuracy(df_fit_arima_mnth_TS_auto_forecast)
```

##  Forecast ARIMA model(kitchen)
```{r Forcast_arima_kit}
#will calculate everything automatic 
df_fit_arima_mnth_kit_auto<-auto.arima(df_arima_Kitchen_stuff)
Box.test(df_fit_arima_mnth_kit_auto$residuals,
         lag=20,
         type="Ljung-Box")
hist(df_fit_arima_mnth_kit_auto$residuals)
#lines(density(df_fit_arima_mnth_TS_auto$residuals))
checkresiduals(df_fit_arima_mnth_kit_auto)

df_fit_arima_mnth_kit_auto_forecast<-forecast(df_fit_arima_mnth_kit_auto,h=12)
autoplot(df_fit_arima_mnth_kit_auto_forecast)
accuracy(df_fit_arima_mnth_kit_auto_forecast)
```


##  Forecast ARIMA model(Laundry)
```{r Forcast_arima_laun}
#will calculate everything automatic 
df_fit_arima_mnth_laun_auto<-auto.arima(df_arima_Laundry)
Box.test(df_fit_arima_mnth_laun_auto$residuals,
         lag=20,
         type="Ljung-Box")
hist(df_fit_arima_mnth_laun_auto$residuals)
#lines(density(df_fit_arima_mnth_TS_auto$residuals))
checkresiduals(df_fit_arima_mnth_laun_auto)

df_fit_arima_mnth_laun_auto_forecast<-forecast(df_fit_arima_mnth_laun_auto,h=12)
autoplot(df_fit_arima_mnth_laun_auto_forecast)
accuracy(df_fit_arima_mnth_laun_auto_forecast)
```

## Fit Linear Regression Model to Monthly Time Series(Water_heater_AC)
```{r Linear Regression_wh_ac}
fit2 <- tslm(housePWR_mnthTS[,3] ~ trend + season)
y <- forecast(fit2,h=12, level=c(80,95))
plot(y, showgap=FALSE, include=36,
  #shadecols=c('slategray3','slategray'),
  xlab ='Year',
  ylab=' kWh',
  main='12-Month Forecast of Monthly Energy Consumption')
minor.tick(nx=6)

accuracy(y)
summary(y)
summary(fit2)

#scatterplot of predicted vs actual
plot.ts(x=fit2$fitted.values, y=housePWR_mnthTS[,3], xy.lines = FALSE,
        xy.labels = FALSE,
        xlab='Fitted Value',
        ylab='Actual',
        main='Monthly Predicted vs. Actual Values')
#abline(0,1, col='blue')

plot.ts(x=fit2$fitted.values, y=fit2$residuals, xy.lines=FALSE,
        xy.labels = FALSE,
        xlab='Predicted Value',
        ylab='Residuals',
        main='Monthly Predicted vs. Residuals')
#abline(0,0, lty=2, col='grey')

qqnorm(fit2$residuals, main='Monthly Q-Q Plot')
#qqline(fit2$residuals)
#-Summary analysis of residuals for monthly linear model
checkresiduals(fit2)

```

## Fit Linear Regression Model to Monthly Time Series(Kitchen)
```{r Linear Regression_kitchen}
fit2_kit <- tslm(housePWR_mnthTS[,1] ~ trend + season)
kit <- forecast(fit2_kit,h=12, level=c(80,95))
plot(kit, showgap=FALSE, include=36,
  #shadecols=c('slategray3','slategray'),
  xlab ='Year',
  ylab=' kWh',
  main='12-Month Forecast of Monthly Energy Consumption(Kitchen)')
minor.tick(nx=6)

accuracy(kit)
summary(kit)
summary(fit2_kit)

#scatterplot of predicted vs actual
plot.ts(x=fit2_kit$fitted.values, y=housePWR_mnthTS[,1], xy.lines = FALSE,
        xy.labels = FALSE,
        xlab='Fitted Value',
        ylab='Actual',
        main='Monthly Predicted vs. Actual Values')
#abline(0,1, col='blue')

plot.ts(x=fit2_kit$fitted.values, y=fit2$residuals, xy.lines=FALSE,
        xy.labels = FALSE,
        xlab='Predicted Value',
        ylab='Residuals',
        main='Monthly Predicted vs. Residuals')
#abline(0,0, lty=2, col='grey')

qqnorm(fit2_kit$residuals, main='Monthly Q-Q Plot')
#qqline(fit2$residuals)
#-Summary analysis of residuals for monthly linear model
checkresiduals(fit2_kit)

```

## Fit Linear Regression Model to Monthly Time Series(Laundry)
```{r Linear Regression_Laundry}
fit2_laun <- tslm(housePWR_mnthTS[,2] ~ trend + season)
laun <- forecast(fit2_laun,h=12, level=c(80,95))
plot(laun, showgap=FALSE, include=36,
  shadecols=c('slategray3','slategray'),
  xlab ='Year',
  ylab=' kWh',
  main='12-Month Forecast of Monthly Energy Consumption(Laundry)')
minor.tick(nx=6)

accuracy(laun)
summary(laun)
summary(fit2_laun)

#scatterplot of predicted vs actual
plot.ts(x=fit2_laun$fitted.values, y=housePWR_mnthTS[,2], xy.lines = FALSE,
        xy.labels = FALSE,
        xlab='Fitted Value',
        ylab='Actual',
        main='Monthly Predicted vs. Actual Values')
#abline(0,1, col='blue')

plot.ts(x=fit2_laun$fitted.values, y=fit2_laun$residuals, xy.lines=FALSE,
        xy.labels = FALSE,
        xlab='Predicted Value',
        ylab='Residuals',
        main='Monthly Predicted vs. Residuals')
#abline(0,0, lty=2, col='grey')

qqnorm(fit2_laun$residuals, main='Monthly Q-Q Plot')
#qqline(fit2$residuals)
#-Summary analysis of residuals for monthly linear model
checkresiduals(fit2_laun)

```
